---
title: "Getting Started"
author: "Christina Maimone"
date: "1/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# RStudio Tour

* What is RStudio
* What are the different panes for?
* Console
  * up arrow for previous commands
  * () for grouping `(3+1)\*4`
  * Call a function: `round(2.1)`, `log(16)`, `round(log(16))`, `round(log(16), 2)`
  * assign a value: `x <- 3` >> shows up in environment
  * vector: `x <- c(1, 4, 3, 2, 5)`
  * history
  * case sensitive
  * spaces don't matter
  * line continuation, escaping
  * Getting stuck with +
* Files
  * `getwd()`, `setwd()`
  * Projects
  * Send code from file to console
    * Send a comment line and it sends the line below
    * Sends a full command, even over multiple lines
  * New Files
    * Different types: R script, R Markdown, R Notebook
    * Comment character
  * Red, \* for files with unsaved changes
* Packages
  * installing
  * loading


Cheatsheet: under Help > Cheatsheets > Browse Cheatsheets

# Do Something!

## Get Data

Import a data set.  Rows are observations, columns are variables.

```{r}
library(readr)  # makes importing data a bit easier

africa <- read_csv("africadata.csv")  # works because this file is in my working directory

# if you haven't downloaded this file and the data file, read from the internet instead:
africa <- read_csv("https://nuitrcs.github.io/mmss211R/africadata.csv")

africa
```

* What are the `NA` values in the rail column?
* What are the numbers to the left of the country names?
* What is a tibble?

```{r, eval=FALSE}
View(africa)
```

## Plot

What is the distribution of country sizes?  Use the area variable (column).

```{r}
hist(africa$area, breaks=15)
```

Or, using `ggplot2` and `dplyr` packages (from the `tidyverse`):

```{r}
library(tidyverse)  # loads multiple packages, including ggplot2 and dplyr

africa %>%
  ggplot(aes(area)) + 
  geom_histogram(bins=15)
```

* What is `%>%`?
* Why is there a `+`?
* What is `aes()`

Why `ggplot2`?

```{r}
africa %>%
  ggplot(aes(area)) + 
  geom_histogram(bins=15) + 
  facet_grid(region~.)
```



## Summarize

```{r}
mean(africa$area)
mean(africa$rail)
mean(africa$rail, na.rm=TRUE)
summary(africa)
```


## Indexing

You can select elements of R objects by index or name.  Index values start at 0:

```{r}
africa[1, "rail"]
africa[1, 6]
africa$rail[1]
```

You can also select with a vector `TRUE` and `FALSE` values the same length as the thing you're selecting from, or expressions that evaluate to `TRUE` and `FALSE`.

```{r}
africa$country == "Sudan"
africa[africa$country == "Sudan", ]
```

Note the `==` for testing for equality, and leaving the index after , empty means to select all columns.

```{r}
africa[africa$road < mean(africa$road), c("country", "road")]  
```

Note to select multiple columns by name, we put them in a vector with `c()`.

## Transformation

We can use vectors (variables) in expressions, and also create new ones:

```{r}
africa$gdp_per_cap <- africa$gdp_2017/africa$pop_2017
africa
```

# Documentation

Not really "help."

Basic functions: https://cran.r-project.org/doc/contrib/Baggott-refcard-v2.pdf

See resources below.


# Try It!

1. Make a scatterplot of GDP vs. population using the africa data.  

2. Which country in Africa has the largest road system (total km of roads)?

3. Which country in Africa has the highest ratio of km of roads to area (area is in square km)?  


# Resources

## DataCamp Courses

In addition to introductory/general R courses on DataCamp, there are a few courses with material that may be relevant to your course:

* Linear regression: https://www.datacamp.com/courses/correlation-and-regression
* Markov: https://www.datacamp.com/community/tutorials/markov-chain-analysis-r
* Networks: https://www.datacamp.com/courses/network-analysis-in-r


## Next Steps

Two free, online books that use the tidyverse set of packages may be helpful to you:

* https://moderndive.com/: the [getting started](https://moderndive.com/1-getting-started.html) chapter can be a good review of things we discussed
* https://r4ds.had.co.nz/ 


## Links to More Resources

Example R Markdown files covering introductory R material: https://github.com/nuitrcs/intro_r_workshop/tree/master/corelecturenotes 

Links to R Resources (and longer workshop materials): https://github.com/nuitrcs/rworkshops 


## Additional Help 

[Drop-in hours](https://www.it.northwestern.edu/research/consultation/data-services.html) - we won't help with your coursework, but you're welcome to come and ask questions about how to use R.