---
title: "Part 2"
author: "Christina Maimone"
date: "1/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Try It Answers

From last time...

1. Make a scatterplot of GDP vs. population using the africa data.  


```{r}
africa <- read.csv("https://nuitrcs.github.io/mmss211R/africadata.csv")
plot(x=africa$pop_2017, y=africa$gdp_2017)
```

Alternate syntax:

```{r, eval=FALSE}
plot(gdp_2017 ~ pop_2017, data=africa)
```



2. Which country in Africa has the largest road system (total km of roads)?

```{r}
max(africa$road)
africa[africa$road == max(africa$road),]
```


3. Which country in Africa has the highest ratio of km of roads to area (area is in square km)?  

```{r}
africa$road/africa$area
max(africa$road/africa$area)
africa$road_ratio <- africa$road/africa$area
africa[africa$road_ratio == max(africa$road_ratio),]
```



# Regression

Function for regression is `lm()`:

```{r}
lm(road ~ pop_2017 + gdp_2017, data=africa)
```

- `road` is the y (dependent) variable.
- `lm()` automatically adds an intercept (constant) term
- `pop_2017` and `gdp_2017` are independent/predictive/x variables

The above doesn't give us much info.  Let's save the output:

```{r}
reg1 <- lm(road ~ pop_2017 + gdp_2017, data=africa)
reg1
```

To get more info, use `summary`:

```{r}
summary(reg1)
```

If you need to access specific pieces of information from the results, there are two places it might be.  First, part of the regression object itself:

```{r}
names(reg1)
reg1$coefficients
```

or part of the results of calling the summary function:

```{r}
names(summary(reg1))
summary(reg1)$r.squared
```

If you plot the regression object, you get diagnostic plots.  Understanding how to interpret these is beyond the scope of this course.

```{r}
plot(reg1)
```




## Formula Syntax

Symbol | Example | Description
-------|-----------|----------------------------------
`~` | `y ~ x1` | Defines the formula (necessary to create a formula object)
`+` | `y ~ x1 + x2` | Include the variable
`-` | `y ~ -1 +x1` | Delete a term, usually a 1 for the intercept
`:` | `y ~ x1 + x1:x2` | Interaction term 
`*` | `y ~ x1*x2` | Interaction between the variables and each individually; same as `y ~ x1 + x2 + x1:x2`
`I()` | `y ~ I(x1^2)` | Wrapper for transforming variables without having to create a new variable

There are additional symbols, but these are all of the ones I've actually used.


```{r}
reg2 <- lm(road ~ pop_2017 * gdp_2017, data=africa)
summary(reg2)
```




# Resources

* DataCamp courses cover regression in R
* A quick run through: https://nuitrcs.github.io/intro_r_workshop/extratopics/basicstats.html
* UCLA stats consulting: https://stats.idre.ucla.edu/other/dae/